---
title: "Meals paper analyses - Directives"
author: "Virginia Marchman (marchman@stanford.edu) and Monica Ellwood-Lowe (mellwoodlowe@berkeley.edu"
date: "April 16, 2022"
output:
  html_document: 
   toc: true
   toc_float: true
---

```{r, echo = F}
library(knitr)
opts_chunk$set(echo=TRUE, 
               warning=FALSE, message=FALSE, 
               cache=FALSE)
options(width = 100)
```

This is the code to generate the analyses in Ellwood-Lowe, M. E., Mastin, J. D., Marchman, V. A., & Fernald, A. (2022) "Big boys eat their peas!": Behavior talk during mealtimes in a diverse sample of English-speaking families.

Load libraries and set theme
```{r}

library(tidyverse)
library(effsize)
library(stargazer)
library(psych)
library(powerAnalysis)
library(kableExtra)
library(knitr)
library(ppcor)

theme_set(theme_bw())
```


# Data prepping
### Load data
```{r}

Meals <- read_csv("./Meals_data_FINAL_forR.csv")

Meals <- as.data.frame(Meals)
```

### Check data - sex, birthorder, age
```{r}

# Sex - 

Meals %>% 
  group_by(sex) %>% 
  count()

# birthorder

Meals_birthorder <- Meals %>%
  mutate(
    birthgroup = case_when(
          birthorder == 1 ~ 1,
          birthorder == 2 ~ 2,
          birthorder > 2 ~ 3)
    )

Meals_birthorder %>% 
  group_by(birthgroup) %>% 
  count()
  
```

### Descriptive statistics - age, maternal education, SES (HI) - Table 1
```{r}


Meals_desc <- dplyr::select(Meals, Age18, Age24, momed, HI18Final)

describe(Meals_desc)

```


```{r}

# Recording lengths

Meals_reclengths <- dplyr::select(Meals, SumDursecAD, SumDurSecCleaned, MEALDUR, NumPplMeal) %>%
                                  mutate(rawlength = as.numeric(SumDursecAD/3600), 
                                         cleanedlength = as.numeric(SumDurSecCleaned/3600),
                                         mealLengthmins = as.numeric(MEALDUR/60))

describe(Meals_reclengths)

```

# Correlation between all day CDS/hour and meals
```{r}

corr.test(Meals$ADCDSperhour, Meals$CDSWCMealperhour)

lm_Meals_SES <- 
  lm(formula = ADCDSperhour ~ CDSWCMealperhour + HI18Final, data = Meals)

lm_Meals_SES_int <- 
  lm(formula = ADCDSperhour ~ CDSWCMealperhour * HI18Final, data = Meals)

summary(lm_Meals_SES)
summary(lm_Meals_SES_int)

anova(lm_Meals_SES, lm_Meals_SES_int)

```

## Descriptives of Talk measures - Table 2

```{r}

## Frequencies

Meals_freqs <- dplyr::select(Meals, ADCDSperhour, CDSWCMealperhour, CDSUttsMealperhour,
                      BTperhour, Directivesperhour, Instrperhour, Prohperhour, Directperhour,
                      Indirperhour, Feedbackperhour,InformFBperhour, UninfFBperhour,
                      PosFBperhour, NegFBperhour)

describe(Meals_freqs)

## Proportions

Meals_props <- dplyr::select(Meals, CDSWCMealper, C_CDS_UTTper, C_BehaviorTalkPER,
                             C_DirectivesPerBT, C_InstructionsPerDIR, C_ProhibitionsPerDIR,
                             C_DirectPerDIR, C_IndirectPerDIR,
                             C_FeedbackPerBT, C_InformFeedbackPerFB, C_NotInformFeedbackPerFB,
                             C_PosFeedbackPerFB, C_NegFeedbackPerFB)
     
describe(Meals_props)

```
```

## Descriptives for MLU-words

```{r}

Meals_MLU <- dplyr::select(Meals, totBTMLU, TotalDirMLU, InstructionsMLU, ProhibMLU,
                          IndirectMLU, DirectMLU, TotFBMLU, InfFBMLU, UninfFBMLU, 
                          PosFBMLU, NegFBMLU)
     
describe(Meals_MLU)

```

```{r}

## paired sample t-tests - Directives

t.test(Meals_MLU$InstructionsMLU, Meals_MLU$ProhibMLU, paired = TRUE, 
  alternative = "two.sided")

t.test(Meals_MLU$IndirectMLU, Meals_MLU$DirectMLU, paired = TRUE, 
  alternative = "two.sided")

## paired sample t-tests - Feedback

t.test(Meals_MLU$InfFBMLU, Meals_MLU$UninfFBMLU, paired = TRUE, 
  alternative = "two.sided")

t.test(Meals_MLU$PosFBMLU, Meals_MLU$NegFBMLU, paired = TRUE, 
  alternative = "two.sided")

```

# Correlations SES (HI) and measures of talk during meals

```{r}
## zero order

cor.test(Meals$HI18Final, Meals$C_CDS_UTTper)
cor.test(Meals$HI18Final, Meals$C_BehaviorTalkPER)
cor.test(Meals$HI18Final, Meals$C_DirectivesPerBT)
cor.test(Meals$HI18Final, Meals$C_InstructionsPerDIR)
cor.test(Meals$HI18Final, Meals$C_ProhibitionsPerDIR)
cor.test(Meals$HI18Final, Meals$C_DirectPerDIR)
cor.test(Meals$HI18Final, Meals$C_IndirectPerDIR)
cor.test(Meals$HI18Final, Meals$C_FeedbackPerBT)
cor.test(Meals$HI18Final, Meals$C_InformFeedbackPerFB)
cor.test(Meals$HI18Final, Meals$C_NotInformFeedbackPerFB)
cor.test(Meals$HI18Final, Meals$C_PosFeedbackPerFB)
cor.test(Meals$HI18Final, Meals$C_NegFeedbackPerFB)

## partial correlations controlling for overall CDS Words/hour

pcor.test(x = Meals$HI18Final, y = Meals$C_CDS_UTTper, z = Meals$CDSWCMealperhour)
pcor.test(x = Meals$HI18Final, y = Meals$C_BehaviorTalkPER, z = Meals$CDSWCMealperhour)
pcor.test(x = Meals$HI18Final, y = Meals$C_DirectivesPerBT, z = Meals$CDSWCMealperhour)
pcor.test(x = Meals$HI18Final, y = Meals$C_InstructionsPerDIR, z = Meals$CDSWCMealperhour)
pcor.test(x = Meals$HI18Final, y = Meals$C_ProhibitionsPerDIR, z = Meals$CDSWCMealperhour)
pcor.test(x = Meals$HI18Final, y = Meals$C_DirectPerDIR, z = Meals$CDSWCMealperhour)
pcor.test(x = Meals$HI18Final, y = Meals$C_IndirectPerDIR, z = Meals$CDSWCMealperhour)
pcor.test(x = Meals$HI18Final, y = Meals$C_FeedbackPerBT, z = Meals$CDSWCMealperhour)

```

```{r}
## must put in new dataframe since 1 person had no feedback

Meals_FB <- filter(Meals, C_InformFeedbackPerFB != "NA")

pcor.test(x = Meals_FB$HI18Final, y = Meals_FB$C_InformFeedbackPerFB, z = Meals_FB$CDSWCMealperhour)
pcor.test(x = Meals_FB$HI18Final, y = Meals_FB$C_NotInformFeedbackPerFB, z = Meals_FB$CDSWCMealperhour)
pcor.test(x = Meals_FB$HI18Final, y = Meals_FB$C_PosFeedbackPerFB, z = Meals_FB$CDSWCMealperhour)
pcor.test(x = Meals_FB$HI18Final, y = Meals_FB$C_NegFeedbackPerFB, z = Meals_FB$CDSWCMealperhour)

```

```{r}

## Plot between proportion behavior talk and SES

theme_set(theme_bw())
ggplot(Meals, mapping = aes(HI18Final, C_BehaviorTalkPER)) + 
  geom_point(color = "black", size = 1.75) + 
  geom_smooth(method = "lm") +
  theme(text = element_text(size = 20)) + 
  theme(panel.grid = element_blank()) +
       # panel.grid.major.y = element_line(colour = "grey50")) +
  labs(x = "Family SES (Hollingshead Index)", y = "Proportion\nBehavior Talk") + 
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) #+ 
  #annotate("text", x = 2, y = 26.5, label = "paste(R ^ 2, \" = .23\")", parse = TRUE, size = 6) # change this R2 value for each plot

ggsave("./scatterplot_behtalkxSES.jpg", height = 8, width = 11, units = "in", dpi = 300)

```

```{r}

#Descriptives for CDI at 18 and 25 months

Meals_CDI <- dplyr::select(Meals, WS18Vocab, WS24Vocab)

describe(Meals_CDI)

```

```{r}

# Vocab correlations with SES

cor.test(Meals$HI18Final, Meals$WS18Vocab)
cor.test(Meals$HI18Final, Meals$WS24Vocab)

```

```{r}

# Zero-order correlations to vocab at 24 months

cor.test(Meals$C_DirectivesPerBT, Meals$WS24Vocab)
cor.test(Meals$C_FeedbackPerBT, Meals$WS24Vocab)

cor.test(Meals$C_InstructionsPerDIR, Meals$WS24Vocab)
cor.test(Meals$C_ProhibitionsPerDIR, Meals$WS24Vocab)

cor.test(Meals$C_IndirectPerDIR, Meals$WS24Vocab)
cor.test(Meals$C_DirectPerDIR, Meals$WS24Vocab)

cor.test(Meals$C_InformFeedbackPerFB, Meals$WS24Vocab)
cor.test(Meals$C_NotInformFeedbackPerFB, Meals$WS24Vocab)

cor.test(Meals$C_PosFeedbackPerFB, Meals$WS24Vocab)
cor.test(Meals$C_NegFeedbackPerFB, Meals$WS24Vocab)


```

# Regression models
###  Table 3
```{r}

# covariates SES, CDS words/hour, Vocab18
m1 <- lm(WS24Vocab ~ HI18Final + CDSWCMealperhour + WS18Vocab, data = Meals)

# Adding Proportion Feedback (vs. Directives) out of all Behavior Talk
m2 <- lm(WS24Vocab ~ HI18Final + CDSWCMealperhour + WS18Vocab +
    C_FeedbackPerBT, data = Meals)

# Adding Proportion Instructive (vs. Prohibitions) out of all Directives
m3 <- lm(WS24Vocab ~ HI18Final + CDSWCMealperhour + WS18Vocab +
    C_InstructionsPerDIR, data = Meals)

# Adding Proportion Indirect (vs. Direct) out of all Directives
m4 <- lm(WS24Vocab ~ HI18Final + CDSWCMealperhour + WS18Vocab + 
    C_IndirectPerDIR, data = Meals)

# comparisons
anova(m1, m2)
anova(m1, m3)
anova(m1, m4)

stargazer(m1, m2, m3, m4, type = "text",
         star.char = c(".","*","**","***"),
         star.cutoffs = c(.1, .05, .01, .001),
         notes = c(". p<0.1; * p<0.05; ** p<0.01; *** p<0.001"),
         notes.append = F,
         digits = 3,
         dep.var.labels = c("Vocabulary Size at 24 months"),
         covariate.labels=c("SES", "CDS Words/hour",
                            "Vocabulary Size at 18 months",
                            "Proportion Feedback",
                            "Proportion Instructive",
                            "Proportion Indirect"))
```

```{r}

# Partial plot
# Proportion Indirect and Vocab at 24m (from Model 4)

## Proportion indirect residuals - not including vocab at 24
m_indirect_resid <- lm(C_IndirectPerDIR ~ HI18Final + CDSWCMealperhour +
  WS18Vocab, data = Meals)
resid_indirect <- resid(m_indirect_resid)

# cdi residuals - not including proportion indirect
m_vocab_resid <- lm(WS24Vocab ~ HI18Final + CDSWCMealperhour +
  WS18Vocab, data = Meals)
resid_vocab <- resid(m_vocab_resid)

# create dataframe
resid_vocab_indirect <- data.frame(resid_indirect, resid_vocab)

# correlation
corr_vocab_indirect <- cor.test(resid_indirect, resid_vocab)
round(((corr_vocab_indirect$estimate)^2), digits = 2)
round(((corr_vocab_indirect$estimate)), digits = 2)

# plot
theme_set(theme_bw())
ggplot(resid_vocab_indirect, aes(resid_indirect, resid_vocab)) + 
  geom_point(color = "black", size = 1.75) + 
  geom_smooth(method = "lm") +
  theme(text = element_text(size = 20)) + 
  theme(panel.grid = element_blank()) +
       # panel.grid.major.y = element_line(colour = "grey50")) +
  labs(x = "Proportion Indirect Directives (residuals)", y = "Vocabulary Size\nat 24 months\n(residuals)") + 
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) #+ 
  #annotate("text", x = 2, y = 26.5, label = "paste(R ^ 2, \" = .23\")", parse = TRUE, size = 6) # change this R2 value for each plot

ggsave("./scatterplot_resid_indirect_vocab.jpg", height = 8, width = 11, units = "in", dpi = 300)

```


